Developing a preprint culture in biology

### 2012-09-28 ms.tex
discussion

Submitting to a preprint server is becoming more common in biology, even though
it still involves a minority of papers. The quantitative biology section in
arXiv is experiencing faster growth in submissions than any other field
\cite{cal12}. Also, most scientific journals are preprint-friendly: Nature,
PLOS, BMC, PNAS, Science (mostly) \ref{table:policies}, and all the journals
from Elsevier and Springer. Very recently, the Ecological Society of America and
the Genetics Society of America changed their policies to allow public
preprints. Few scientific publications will not consider a manuscript submitted
to a public preprint server.  Still, a few journals adopt a ``by default''
hostile attitude towards preprints, mostly due to the lack of clear policy of
the publishers, or perhaps because a preprint culture has not developed in
biology and the practice is still considered unusual. As an example,
Wiley-Blackwell, which publishes some leading journals, has no official policy
on the subject \ref{table:policies}.

@@

Figshare (\href{http://figshare.com/}{http://figshare.com/}) is an open server
that allows scientists to submit any research output: manuscript, figures,
datasets, videos, theses, presentations, and so on. There are no rules to limit
what constitutes a research output and, unlike arXiv, there is no endorser
system. All figshare content has a unique digital object identifier (DOI) like
any journal article, thus offering a permanent and stable link to the content.
A flexible tag system is used to classify each item. All content can be
commented and is licensed under the Creative Commons (CC-BY) license, except
datasets which are published under CC0.

One of the biggest advantage of figshare over arXiv is that is it not limited to
quantitative sciences. arXiv.org has sections on quantitative biology but might
not be appropriate for non-quantitative work. With its flexible approach to
preprints, figshare offer an important alternative to arXiv for empirical
biologists. Furthermore, by allowing all

@@

Responding to the rumour that they refused manuscripts submitted to arXiv,
Nature responded that ``Nature never wishes to stand in the way of communication
between researchers. We seek rather to add value for authors and the community
at large in our peer review, selection and editing'' \cite{nat05}.

Open preprints server offer a great opportunity for open science, especially if
the community embrace the idea of discussing preprints. Initiatives like
Haldane's Sieve (\href{http://haldanessieve.org/}{http://haldanessieve.org/}), a
new blog discussing arXiv papers in population genetics, will help make arXiv
attractive for scientists looking to promote their work. These initiatives are
important to fully exploit the potential of open preprints servers.

Posting preprints online increases the community of available informal peer
reviewers, and uses the internet for its original community-building purposes.
Preprint servers also facilitate communication between disciplines, bridging
cultural as well as geographic divides. The advantages are clear and the costs
are low.

Physicists and other quantitative scientists have recently developed a great
interest in evolutionary theory, quantitative ecological theory and
epidemiological modeling.  Unfortunately this has led to a lot of repeated work
(``reinventing the wheel''; \cite{de2011contribution}) that could be avoided by
better communication across disciplines.  The near-universal adoption of
preprint servers in physics provides this vital communication channel.  We
suggest that biologists reach out to physicists and mathematicians by posting
papers to arXiv: a physicist who does not read \emph{Evolution} certainly checks
arXiv at least weekly.  This benefits both disciplines, as biologists will reach
new readers, and physicists will learn the terminology, tools and ideas common
in existing evolutionary theory.  A recent series of papers on the theory of
natural selection was posted to arXiv simultaneously with its publication in the
\emph{Journal of Evolutionary Biology}
\cite{JEB:JEB2431,JEB:JEB2498,JEB:JEB2378,JEB:JEB2373}.

>>>>

Submitting to a preprint server is becoming more common in biology, even though
it still involves a minority of papers. The quantitative biology section in
arXiv is experiencing faster growth in submissions than any other field
\cite{cal12}. Examples include a recent series of papers on the theory of
natural selection that was posted to arXiv simultaneously with its publication
in the \emph{Journal of Evolutionary Biology}
\cite{JEB:JEB2431,JEB:JEB2498,JEB:JEB2378,JEB:JEB2373}.

Most scientific journals are preprint-friendly: Nature, PLOS, BMC, PNAS, Science
(mostly) \ref{table:policies}, and all the journals from Elsevier and Springer.
Very recently, the Ecological Society of America and the Genetics Society of
America changed their policies to allow public preprints. Few scientific
publications will not consider a manuscript submitted to a public preprint
server.  Still, a few journals adopt a ``by default'' hostile attitude towards
preprints, mostly due to the lack of clear policy of the publishers, or perhaps
because a preprint culture has not developed in biology and the practice is
still considered unusual. As an example, Wiley-Blackwell, which publishes some
leading journals, has no official policy on the subject \ref{table:policies}.

@@

Figshare (\href{http://figshare.com/}{http://figshare.com/}) is an open server
that allows scientists to submit any research output: manuscript, figures,
datasets, videos, theses, presentations, and so on. There are no rules to limit
what constitutes a research output and, unlike arXiv, there is no endorser
system. All figshare content has a unique digital object identifier (DOI) like
any journal article, thus offering a permanent and stable link to the content.
A flexible tag system is used to classify each item. All content can be
commented and is licensed under the Creative Commons (CC-BY) license, except
datasets which are published under CC0. The CC-BY license grant the rights to
\emph{copy, distribute, display and perform the work and make derivative works
based on it only if they give the author or licensor the credits in the manner
specified by these.} The CC0 is the most permissive license and make the work
public domain (no rights reserved) or, if it is not possible in the given
jurisdiction, provides a simple permissive license.

One of the biggest advantage of figshare over arXiv is that is it not limited to
quantitative sciences. arXiv.org has sections on quantitative biology but might
not be appropriate for non-quantitative work. With its flexible approach to
preprints, figshare offer an important alternative to arXiv for empirical
biologists. Furthermore, by allowing all types of contents, figshare arguably
provides an archive for early results (e.g.: figures, lab presentations).

@@

Open preprints servers offer a great opportunity for open science, especially if
the community embrace the idea of discussing preprints. Initiatives like
Haldane's Sieve (\href{http://haldanessieve.org/}{http://haldanessieve.org/}), a
new blog discussing arXiv papers in population genetics, will help make arXiv
attractive for scientists looking to promote their work. These initiatives are
important to fully exploit the potential of open preprints servers. Posting
preprints online increases the community of available informal peer reviewers,
and uses the internet for its original community-building purposes.  Preprint
servers also facilitate communication between disciplines, bridging cultural as
well as geographic divides. The advantages are clear and the costs are low.
Responding to the rumour that they refused manuscripts submitted to arXiv,
Nature responded that ``Nature never wishes to stand in the way of communication
between researchers. We seek rather to add value for authors and the community
at large in our peer review, selection and editing'' \cite{nat05}.
